---
title: "results joint experiment"
output:
  html_document:
    df_print: paged
---

Load data

```{r, warning=FALSE, message=FALSE}
library(here)
library(ggforce)
source(here("R", "analysis-utils.R"))
```

```{r, data}
# exp.name = "toy-blocks-pilot-2"
exp.name = "blockwords"
DATA = load_exp_data(exp.name, use_filtered=FALSE)
```

## Meta info about participants

```{r, meta, echo=FALSE}
df <- DATA$info %>% group_by(prolific_id) %>% distinct() %>%
  mutate(education=as_factor(education), gender=as_factor(gender))
p <- df %>%
  ggplot(aes(x=factor(0), y=timeSpent)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=16, width=0.2) +
  geom_hline(aes(yintercept = mean(df$timeSpent) + sd(df$timeSpent),
                 color="yellow")) +
  geom_hline(aes(yintercept = mean(df$timeSpent) + 2 * sd(df$timeSpent),
                 color="red")) +
  theme(legend.position="none") +
  labs(x = "", y="timeSpent in min.")
p

df %>% ungroup() %>%  group_by(prolific_id) %>%  summary()
```

## Reaction Times per stimulus

```{r, reaction-times, echo=FALSE}
df <- DATA$production %>% dplyr::select(prolific_id, id, RT) %>% group_by(id) %>%
  distinct()

TRIALS = df$id %>% unique()

dat.RT = df %>% group_by(id) %>% summarise(mu=mean(RT), sd=sd(RT), .groups="drop_last")
df <- left_join(df, dat.RT, by="id")
for(pa in seq(1,2)) {
  p <- df %>%  
    ggplot(aes(y=RT)) + geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(x=0, color=prolific_id), width = 0.1, alpha = 0.5) +
    geom_hline(aes(yintercept = mu), color="green", linetype="dashed") +
    theme(legend.position="right", axis.ticks.x = element_blank(), axis.text.x = element_blank()) +
    labs(x="", y="RT in ms", title="Reaction Times per stimulus") +
    facet_wrap_paginate(~id, nrow=2, ncol=5, page=pa) +
    scale_y_continuous(trans = log2_trans(),
                       breaks = trans_breaks("log2", function(x) 2^x),
                       labels = trans_format("log2", math_format(2^.x))) +
    theme(legend.position="none")
  print(p)
}
```

# General comments all participants

```{r, echo=FALSE}
dat <- readRDS(here("data", "prolific", exp.name, paste(exp.name, "_tidy.rds", sep="")));
```

```{r}
df.comments = dat$comments %>% filter(!is.na(comments) & comments!="")
print(df.comments$comments)
```

After looking at comments, save which participants must be excluded due to 
what they said.

```{r}
out.comments = df.comments %>%
  filter(comments == "I apologize because the first couple of graphics did not appear properly, so I picked any answer. The rest of the study went fine." |
         comments == "The descriptions on this quiz were terrible and confusing. Non native English speakers will have a lot of trouble." |
         comments == "This was hard to understand")

# check time they spent on experiment
out.comments_time = 
  left_join(DATA$info %>% filter(prolific_id %in% out.comments$prolific_id),
          out.comments, by = "prolific_id")

out.comments_time

out.comments = out.comments_time %>% dplyr::select(prolific_id)
# remove all trials for these participants
df.out = out.comments %>% 
  add_column(id=rep(list(TRIALS), nrow(out.comments))) %>%
  unnest(c(id)) %>% group_by(prolific_id)
write_csv(df.out, paste(DATA$result_dir, "out_by_comments.csv", sep=fs))
```


# Color vision trials all participants

```{r, color-vision, echo=FALSE}
df <- dat$color %>% group_by(prolific_id, id) %>%
  mutate(correct = expected == response)

df %>%
  ggplot(aes(x=id)) + geom_bar(position="dodge", aes(fill=correct)) +
  theme(axis.text.x = element_text(angle=90, hjust=1), legend.position = "top")
pids.false_colors = df %>% filter(correct == FALSE)  %>% pull(prolific_id) 
```

Attention check questions

```{r}
dat <- readRDS(paste(DATA$result_dir, fs, exp.name, "_tidy.rds", sep=""));
df.att = dat$train.attention %>% filter(response != expected) 
df.att
```



## Production Task
Custom responses

```{r, custom-responses, fig.height = 10, fig.width = 10}
dat.custom = DATA$production %>% filter(!is.na(custom_response)) %>%
  group_by(prolific_id, id) %>% mutate(custom_response = tolower(custom_response))
total = DATA$production %>% dplyr::select(prolific_id, id) %>% distinct()
print(paste('nb total custom responses given:', nrow(dat.custom), 'out of', nrow(total), 
            '(', nrow(dat.custom)/nrow(total), ')'))
df = dat.custom %>% rename(res.given=response, response=custom_response) %>%
  standardize_sentences() %>%
  mutate(newUtt = !response %in% standardized.sentences) %>%
  filter(newUtt) %>% rename(custom_response=response)

df %>% group_by(custom_response) %>% mutate(n=n()) %>% 
  ggplot(aes(x=n, y=fct_rev(fct_infreq(custom_response)))) +
  geom_bar(aes(fill=id), stat="identity", position=position_dodge()) +
  geom_text(aes(label=res.given), position = position_nudge(x=-0.5, y=-0.25)) +
  theme(text = element_text(size=20),
        axis.text.x=element_text(angle=45, vjust = 0.5),
        legend.position="none") +
  labs(x="#participants", y="custom response", title="(not creatable) custom responses") +
  theme_bw()
```

For each stimulus, which response was created how often?

```{r, echo=FALSE, fig.height = 12}
df.production.means = DATA$production %>%
  dplyr::select(id, response, prolific_id) %>% 
  group_by(id,response) %>% 
  mutate(n=n()) %>% group_by(id) %>% mutate(N=n(), ratio=n/N) %>%
  arrange(desc(ratio)) %>% 
  add_column(predictor="empirical")

df = df.production.means %>% dplyr::select(-prolific_id, -n, -N) %>% distinct() %>% 
  mutate(response=as_factor(response)) %>% 
  filter(ratio>0)

df %>% 
  ggplot(aes(y=response, x=ratio)) +
  geom_bar(aes(fill=id), stat="identity") +
  theme_bw(base_size=18) +
  theme(legend.position = "none") +
  facet_wrap(~id)
```



```{r}
df.production.means = DATA$production %>% filter(id != "ind2") %>%
  dplyr::select(response, prolific_id, id) %>% 
  group_by(prolific_id,response) %>% 
  mutate(n=n()) %>% group_by(prolific_id) %>% mutate(N=n(), ratio=n/N) %>%
  arrange(desc(ratio)) %>% distinct() %>%
  mutate(response=as.factor(response))

# also note time spent
df = left_join(df.production.means,
               DATA$info %>% dplyr::select(prolific_id, timeSpent), 
               by=c("prolific_id")) %>% 
  mutate(timeSpent=round(timeSpent, 2))

df$pid = df %>% group_by(prolific_id) %>% 
  group_indices()

p  = df %>%
  dplyr::select(pid, response, n, timeSpent) %>% distinct() %>% 
  ggplot() +
  geom_bar(aes(y=response, x=n), stat="identity") +
  geom_text(aes(x=9, y="green might not fall", label=timeSpent), color="red") + 
  theme_bw(base_size=18) +
  facet_wrap(~pid) + 
  theme(legend.position = "none") +
  ggtitle("utterance frequencies per participant with total time spent in red")
  
ggsave(paste(DATA$plot_dir, "utterance-frequencies.png", sep=fs),
       p, width=18, height=22)


# participants who choose <= 3 different utterances AND whose total time spent
# was less than 20 minutes
df.sum = df %>%
  dplyr::select(response, prolific_id, pid, timeSpent) %>% distinct() %>%
  group_by(prolific_id) %>% mutate(n.utt=n()) %>%
  dplyr::select(-response) %>% distinct()

df.out = df.sum %>% 
  filter(n.utt <= 3 & timeSpent < 20) 
df.out

# remove all trials for these participants
df.out = df.out %>% dplyr::select(prolific_id) %>%
  add_column(id=rep(list(TRIALS), nrow(df.out))) %>%
  unnest(c(id)) %>% group_by(prolific_id)
```

# Participants' slider rating for u in task 1 is 0 but u is produced in task 2

```{r}
DATA$joint.orig %>%
  filter(!is.na(human_exp2) & human_exp1 == 0)
```

Data Quality 

```{r}
dat.quality = readRDS(paste(here("data", "prolific", exp.name), fs,
                            "test-data-prior-quality.rds", sep=""))
quality.means = dat.quality %>% arrange(desc(mean.comparator)) %>%
  distinct_at(vars(c(comparator)), .keep_all = TRUE)

p = quality.means %>% 
  ggplot(aes(x=mean.comparator)) +
  geom_density() +
  geom_jitter(aes(y=0, color=comparator), width=0, height=0.2) +
  ggtitle("avg data quality each proband across stimuli") +
  theme(legend.position="none")
p

df.qual = quality.means %>% arrange(desc(mean.comparator))
pids.qual = df.qual[1:3, ]$comparator
qual.time = DATA$info %>% filter(prolific_id %in% pids.qual) %>%
  dplyr::select(prolific_id, timeSpent)
qual.time
# remove
df.out = qual.time %>%
  add_column(id=rep(list(TRIALS), nrow(qual.time))) %>%
  unnest(c(id)) %>% group_by(prolific_id) %>%
  filter(id != "ind2")
write_csv(df.out, paste(DATA$result_dir, "out_by_quality_time.csv", sep=fs))

p = dat.quality %>% distinct_at(vars(c(id)), .keep_all = TRUE) %>% 
  ggplot(aes(x=mean.id)) +
  geom_density() +
  geom_jitter(aes(y=0, color=id), width=0, height=0.2) +
  theme(legend.position="bottom") +
  ggtitle("avg data quality values across stimuili")
p
```

# Training Data

Slider-Choice Trials

```{r}
dat.tidy = readRDS(paste(DATA$result_dir, fs, exp.name, "_tidy.rds", sep=""))
data.train.sliders = dat.tidy$train.slider_choice
data.train.smooth = dat.tidy$train.smooth
data.train.orig = dat.tidy$train.orig
# for each participant only the last 50% of all train trials
data.train.smooth.half = data.train.smooth %>%
  separate(id, into=c("trial.relation", "trial.idx"), sep=-1, remove=FALSE) %>%
  group_by(prolific_id, trial.relation) %>% arrange(desc(trial_number)) %>%
  top_frac(0.5, trial_number) %>%
  dplyr::select(-expected) %>%
  pivot_wider(names_from=question, values_from=response) %>%
  ungroup() %>%
  dplyr::select(-trial.relation, -trial.idx)
  
dat.sliders = data.train.sliders %>%
  mutate(correct = response == expected, .groups="drop_last")
  
# ratio correct per participant
dat.sl.participants = dat.sliders %>% group_by(prolific_id) %>%
  summarize(ratio_correct=sum(correct)/n(), .groups="drop_last")

p = dat.sl.participants %>% 
  ggplot(aes(x=0, y=ratio_correct)) +
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(color=prolific_id), shape=16, width=0.1, height=0) +
  theme(legend.position = "none")
p
dat.sl.participants %>% filter(ratio_correct < 0.5)

# ratio correct per id (difficulty slider-choice trials)
df.sliders = dat.sliders %>% group_by(id) %>% 
  summarize(ratio_correct = sum(correct)/n(), .groups="drop_last") %>%
  arrange(desc(ratio_correct)) %>%
  mutate(id=as_factor(id))
  
df.sliders %>%
  ggplot(aes(y=id, x=ratio_correct)) +
  geom_point() +
  labs(x="ratio participants who gave correct response")
```  

# Training animations
(todo: check if training and prior conditions are matched correctly here! (read from csv))

```{r}
df.train.edges = data.train.smooth.half %>% filter(id %in% train.edges$id) %>%
  group_by(prolific_id, id) %>% 
  mutate(red=ry + r, yellow=ry + y) %>% 
  left_join(train.edges, join_by = c("id")) %>%
  pivot_longer(c(red, yellow, ry, none, r, y), names_to="event", values_to="response")  %>%
  filter(event == block)

df.train.ramp = data.train.smooth.half %>% filter(id %in% train.ramp$id) %>%
  group_by(prolific_id, id) %>% 
  mutate(red=ry + r, yellow=ry + y) %>% 
  left_join(train.ramp, join_by = c("id")) %>%
  pivot_longer(c(red, yellow, ry, none, r, y), names_to="event", values_to="response")  %>%
  filter(event == block)

p = df.train.edges %>%
  mutate(relation=case_when(id %in% c("ssw0", "ssw1") ~ "if2",
                            id %in% c("ac0", "ac1", "ac2", "ac3") ~ "if1",
                            TRUE ~ "independent"),
         condition=factor(condition, levels = c("low", "uncertain", "uncertainH", "high"))) %>% 
  plot_ratings_across_conditions("training-trials block on edge")
p

p = df.train.ramp %>% 
  mutate(relation=case_when(id %in% c("ssw0", "ssw1") ~ "if2",
                            id %in% c("ac0", "ac1", "ac2", "ac3") ~ "if1",
                            TRUE ~ "independent"),
         condition=factor(condition, levels = c("low", "uncertain", "uncertainH", "high"))) %>% 
  plot_ratings_across_conditions("training trials ramp with ball")
p

## Training data: ratings per stimulus, block on edge (low/uncertain/high prior)
p1 = df.train.edges %>% filter(condition=="high") %>% 
  mutate(condition=as.factor(condition)) %>% 
  ggplot(aes(x=condition, y=response)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color=dir), shape=16, width=0.2, alpha=0.5) +
  facet_wrap(~id) +
  theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(size=15)) +
  ggtitle("block on edge") +
  coord_flip()
ggsave(paste(DATA$plot_dir, "train-edges-high.png", sep=.Platform$file.sep), p1, width=15, height=10)
p1

p2 = df.train.edges %>% filter(startsWith(condition, "uncertain")) %>% 
  mutate(condition=as.factor(condition)) %>% 
  ggplot(aes(x=condition, y=response)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color=dir), shape=16, width=0.2, alpha=0.5) +
  facet_wrap(~id) +
  theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(size=15)) +
  ggtitle("block on edge") +
  coord_flip()
ggsave(paste(DATA$plot_dir, "train-edges-uncertain.png", sep=.Platform$file.sep), p2, width=15, height=10)
p2   

p3 = df.train.edges %>% filter(condition == "low") %>% 
  mutate(condition=as.factor(condition)) %>% 
  ggplot(aes(x=condition, y=response)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color=dir), shape=16, width=0.2, alpha=0.5) +
  facet_wrap(~id)+
  theme(axis.text.x = element_text(size=15)) +
  ggtitle("block on edge") +
  coord_flip()
# theme(axis.text.x=element_text(angle=90))
ggsave(paste(DATA$plot_dir, "train-edges-low.png", sep=.Platform$file.sep), p3, width=15, height=10)
p3 
```




